
\clearpage
# Hierarchical Bayesian Model (HBM) of sockeye salmon stock-recruit data in the Skeena and Nass river systems {#app:HBMFits}

```{r,echo=FALSE, message=FALSE}
library(tidyverse)
library(csasdown)

HBM.results <- readRDS("data/HBM/HBM_appendix_tables.rds")
# str(HBM.results)

# TASKS
# - Standardize terms: lognormal or log-normal,  capitalize Normal?
# - methods - include JAGS
# -[ ] intro to senstivity
# - [ ] Citations
# - [ ] shrinkage 
```

<!-- 
Questions

Normal draws were defined in terms of precision, does the remainder of the document

See equation \@ref(eq:HBMLinRick)
-->



The hierarchical Bayesian model (HBM) that was developed for the estimation of Ricker stock-recruit model parameters for sockeye salmon stocks in the Skeena and Nass river systems with independent implementations for the Skeena and Nass River basins.  This structuring was used because it was assumed the exchangeability in the stock productivity parameter was restricted to stocks within each basins.  Differences in the frequency distribution of life history attributes of stocks such as age composition of spawners, the relative frequency of lake rearing and non-lake rearing life history types and differences in run timings and marine migration routes between sockeye salmon stocks in these two watersheds.  

## Mathematical and Statistical Formulation of the HBM {#app:HBMFits-1}

Both the Skeena and Nass HBM models use a linearized form of the Ricker stock-recruit model:


\begin{equation}
\log{\left(\frac{R_{s,y}}{S_{s,y}}\right)=a_s-}\beta_s\cdot\ S_{s,y}+\epsilon_{s,y}
(\#eq:HBMLinRick)
\end{equation}

where $R_{s,y}$ is the observed abundance of recruits produced in stock $s$ in brood year $y$ by the corresponding abundance of spawners brood year $S_{s,y}$. The parameter $a_s$ is the natural logarithm of the maximum rate of population increase, $\beta_s$ is the stock-specific coefficient for the density-dependent effect of spawner abundance on stock productivity and $\epsilon_{s,y}$ represents stock by  year error term that is assumed to be Normally distributed. 


The HBM framework presented is based on that formulated by @KormanEnglish2013 who estimated stock-recruit parameters for numerous salmon stocks in the Skeena River system.  The HBM presumes, as @KormanEnglish2013 did, that the Ricker a parameter is exchangeable between stocks within a river system with a hyper prior mean and hyper prior standard deviation:

\begin{equation}
  a_s \sim  \ln \textrm{Normal}\left(\mu_a,\ \tau_a\right) 
\end{equation}

where $\mu_a$ is the natural logarithm of the hyper prior median of the Ricker $a$ parameter, and $\tau_a$ is the hyperprior precision of the Ricker $a$ parameter.  The following hyperpriors were assigned for $\mu_a$ and $\tau_a$:
  
\begin{equation}
  \mu_a \sim \textrm{Normal}\left({0.5,\ 10^{-6}}\right)
  \textrm{, and }
  \tau_a \sim  \textrm{Gamma} \left(0.5,\ 0.5 \right).
\end{equation}

@KormanEnglish2013 choose these priors as vague priors and the same hyperpriors were retained in the current formulation.

An informative log-normal prior was also assumed for density-dependency parameter  $\beta_s$ which follows the formulation used in @KormanEnglish2013:  
  
\begin{equation}
  \beta_s \sim \ln \textrm{Normal}\left(\mu_{\beta,s},\ \tau_{\beta,s}\right)
\end{equation}

where $\sigma_{max,s}$  is the prior standard deviation in the natural logarithm of $S_{\textrm{max},s}$.  Note that values for $\sigma_{\textrm{max},s}$ are nearly identical to the coefficient of variation (CV) for smaller values of $\sigma_{\textrm{max},s}$ (i.e., about 0.5 and lower) and as such we use CV and $\sigma_{\textrm{max},s}$ interchangeably when describing precision of the prior. That said, the ratio of CV to $\sigma_{\textrm{max},s}$ increases to noticeably larger than 1 (i.e., CV is larger than $\sigma_{\textrm{max},s}$) at larger values of $\sigma_{\textrm{max},s}$ meaning the true CV is larger than stated and the prior will be more diffuse. The values  of $S_{\textrm{max},s}$ and $\sigma_{\textrm{max},s}$ for each stock in the Skeena and Nass watersheds are provided in Table \@ref(tab:PrSmax).



<!-- TABLE 1. Smax Priors -->
```{r PrSmax, echo=FALSE, results='asis'}
cap <- "Prior median values for stock-specific $S_\\textrm{max}$ and the prior standard deviation in the natural logarithm of $S_\\textrm{max}$ (i.e., $\\sigma_{\\textrm{max},s}$) based on results from previous lake productivity analyses.  The two instances where three stocks are listed in the same line, the available mean $S_\\textrm{max}$ values were added since the individual rearing lakes for these stocks were geographically very close together and the stock-recruit data for these stocks were thus combined into a single time series of stock-recruit data for stock-recruit parameter estimation."
# data.frame(Yes = 1:10, No = LETTERS[11:20]) %>%
HBM.results$Smax %>%
  mutate(prSmax = round(prSmax)) %>%
  rename(
    `$S_{\\textrm{max},s}$` = prSmax,
    `CV/$\\sigma_{\\textrm{max},s}$` = prCV
  ) %>%
  csas_table(
    format = "latex",
    escape = FALSE, 
    font_size = 10,
    align = c("l","l", "l", "r", "r"),
    caption = cap
  )
```



A log-normal prior was also assumed for density-dependency parameter  $\beta_s$ which follows the formulation used in @KormanEnglish2013:  

\begin{equation}
  \beta_s \sim \ln \textrm{Normal}\left(\mu_{\beta,s},\ \tau_{\beta,s}\right)
\end{equation}

where $\mu_{\beta,s}$  was obtained based on results from previous limnological lake productivity analyses that assessed the spawner abundance (i.e., $S_{\textrm{max},s})$ which could generate on average the maximum number of recruits for each stock. The values $\mu_{\beta,s}$  and $\tau_{\beta,s}$ were determined as:

  
\begin{equation}
  \mu_{\beta,s}= \log \left(  \frac{1}{S_{\textrm{max},s}} \right)
  \textrm{,  and  } 
  \tau_{\beta,s}=\left(\frac{1}{\sigma_{\textrm{max},s}}\right)^2.
\end{equation}
  
where $\sigma_{\textrm{max},s}$  is the prior standard deviation in the natural logarithm of $S_{\textrm{max},s}$.  Note that values for $\sigma_{\textrm{max},s}$  are nearly identical to the coefficient of variation (CV) for smaller values of $\sigma_{\textrm{max},s}$ (i.e., about 0.5 and lower) and as such we use CV and $\sigma_{\textrm{max},s}$ interchangeably when describing precision of the prior. That said, the ratio of CV to $\sigma_{\textrm{max},s}$ increases to noticeably larger than 1 (i.e., CV is larger than $\sigma_{\textrm{max},s}$) at larger values of $\sigma_{\textrm{max},s}$ meaning the true CV is larger than stated and the prior will be more diffuse. The values  of $S_{\textrm{max},s}$ and $\sigma_{\textrm{max},s}$ for each stock in the Skeena and Nass watersheds are provided in Table \@ref(tab:PrSmax).


The values for $\sigma_{\textrm{max},s}$ were obtained for most of the Skeena sockeye salmon stocks from @KormanEnglish2013. It should be noted that a value for $\sigma_{\textrm{max},s}$ of 2.00 makes the corresponding prior for $S_{\textrm{max},s}$ vague and diffuse. For example,  relative to the median, the lower and upper 95% credible intervals would be about fifty times lower and fifty times higher than the median.  In cases where $\sigma_{\textrm{max},s}$ was set at 2.00, at least some members of the Working Group had raised concerns about the empirical basis for the values for $S_{\textrm{max},s}$ that could be obtained for those stocks.  

No lake productivity estimate was available for Asitka.  A crude approximation of the prior median $S_{\textrm{max},s}$ for Asitka was obtained from the product of the lake surface area for Asitka and the average $S_{\textrm{max},s}$ per unit lake area for Skeena watershed lakes for which S_{max,s} estimates for sockeye salmon were available.  Due the large uncertainty associated with $S_{\textrm{max},s}$ for Asitka, a value of 2.00 was assumed for $\sigma_{\textrm{max},s}$ for Askita which makes this into a vague prior.   

For the five Babine Lake stocks there was admittedly highly uncertain over the appropriate stock-specific lake capacity due to uncertainty over how the common resource is shared between the five stocks as juveniles for at least five of the recognized Babine stocks rear in Babine Lake.  The prior median for $S_{\textrm{max},s}$ for each of the five Babine Lake stocks were obtained by dividing the total Babine lake productivity estimate by  five, providing an equal allocation of the shared resource to each Babine stock.  Because the true proportion of lake productivity used by each stock remains unknown a high level of uncertainty (i.e., a $\sigma_{\textrm{max},s}$  value 2.00) was assigned for each for each Babine Lake stock.

As in @KormanEnglish2013 a Normal likelihood function was applied in the HBM and the same prior was applied to the standard deviation in the deviations between predicted and observed productivity:

\begin{equation}
  \log{\left(\frac{R_{s,y}}{S_{s,y}}\right) \sim   \textrm{Normal}\left(\mu_{s,y}^{RS},\  \tau_s\right)}\ 
\end{equation}


where $\log(R_{s,y}/S_{s,y})$ is the observed natural logarithm of the ratio of recruits produced by spawners (i.e., productivity) in brood year $y$ for stock $s$, $\mu_{s,y}^{RS}$ is the stock-specific productivity predicted by the Ricker stock-recruit model (see equation \@ref(eq:HBMLinRick)), conditioned on parameters $a_s$ and $\beta_s$, and $\tau_s$ is the stock-specific precision used in the likelihood function and is given by:

\begin{equation}
  \tau_s=\frac{1}{\sigma_s^2}
\end{equation}
  
where $\sigma_s$ is the stock-specific standard deviation in the deviates between observed and predicted productivity.  The prior for $\sigma_s$ is the same as that applied by @KormanEnglish2013:

\begin{equation}
  \sigma_s \sim \textrm{Uniform}(0.05,\ 10).
\end{equation}

The original @KormanEnglish2013 HBM did not consider time varying changes in productivity; because this was of interest in the current investigation, the @KormanEnglish2013 HBM was extended to include a common shared year effect on stock productivity (i.e., $\log(R_{s,y}/S_{s,y})$). It is hypothesized that due to the spatial proximity of the sockeye salmon stocks within a basin (i.e., Skeena or Nass basins) and overlap between stocks in migration pathways in freshwater and the marine environment that the annual productivity of different stocks will deviate from Ricker model predictions in the same direction and with similar magnitude in each year.   Therefore, in addition to a stock-specific deviate between the Ricker model prediction of productivity and the observed productivity, a common shared deviate or year effect was also included in the HBM.  The linearized form of the Ricker model that includes the common shared year effect $T_y$ in year $y$ is thus given by:

\begin{equation}
  \mu_{s,y}^{RS}=a_s-\beta_s\cdot S_{s,y}+T_y
\end{equation}
  

The prior for $T_y$ is given by,

\begin{equation}
  T_y \sim \textrm{Normal}\left(0,\ \tau_T\right)
\end{equation}
  
where $\tau_T$  is the prior precision in the common shared year effect that is computed as $\tau_T=1/\sigma_T^2$ with $\sigma_T$ representing the standard deviation in common shared year effects.  The hyperprior for $\sigma_T$  was determined as:

\begin{equation}
  \sigma_T \sim \ln \textrm{Normal} \left(\log{\left(0.8\right),\ \frac{1}{{0.7}^2}}\right).
\end{equation}
  
The prior for $\sigma_T$ was chosen to be mildly informative with the prior central tendency and prior precision set at values to allow the data to inform a large range of values and accommodate potentially large interannual variation in common shared year effects. 

## Model Fitting

The MCMC algorithms applied for posterior integration of the HBM included WinBUGS 1.4.3 and JAGS and practically identical parameter estimates were obtained between the two software packages.    Under initial implementations the burn-in was achieved within about 10,000 iterations and well-pronounced posterior density functions were obtained for parameter estimates.    However, under both software packages, for some of the stocks, the MCMC chains could stray very rarely to extreme low values for the Ricker $\beta$ parameter or extreme high values for the Ricker a parameter.  Extreme low values for the Ricker $\beta$ parameter map out to extreme high values, e.g., ten times the value of the posterior mode, for derived parameters such as the average unfished spawner abundance, $S_0$, or spawning stock abundance associated with the maximum sustainable yield, $S_{msy}$.  Such extreme outlier values were considered to be well outside of the support of the data and may be artifacts of the operation of the MCMC algorithms applied. 

To prevent the inclusion of extreme outlier values, a minimum boundary was applied to the prior for the Ricker $\beta$ parameter.  The prior minimum value for the Ricker $\beta$ parameter was established as follows.  An upper bound on the value for $S_\textrm{max}$ was obtained for each stock as five times the prior central tendency specified for $S_\textrm{max}$:  

\begin{equation}
  S_{\textrm{max},\textrm{max},s} = 5 \times S_{\textrm{max},s}
  \ \textrm{ and }\ \
  \beta_{\textrm{min},s}=\frac{1}{S_{\textrm{max},\textrm{max},s}}. 
\end{equation}

The adjusted prior for $\beta_s$ that was applied was thus:

\begin{equation}
  \textrm{max}\left(\beta_s \sim \ \ln \textrm{Normal}\left(\mu_{\beta,s},\ \ \tau_{\beta,s}\right),\ \beta_{\textrm{min},s}\right)
\end{equation}

The application of this modified prior for $\beta_s$ thus prevented anomalously low values for $\beta_s$ and anomalously large MCMC values for abundance reference points such as  $S_\textrm{max}$, $S_\textrm{msy}$, and $S_0$. 


The posterior predictive distribution for the Ricker $a$ parameter for an unsampled population, $a_p$, is given by:

\begin{equation}
  a_p \sim\  \ln \textrm{Normal}\left(\mu_a,\ \tau_a\right)
  (\#eq:unsamp)
\end{equation}
  
This distribution reflects the effective prior density function for the Ricker $a$ parameter that was applied to estimate this parameter for each stock. 

To provide approximations of how productivity could be varying systematically over years, representations of time varying productivity were obtained by adding the running $n = 4$ or 5 year averages of the common shared year effects to the $a$ parameter for each stock:

\begin{equation}
  a_{s,y,n}=a_s+\frac{1}{n}\cdot\sum_{i=y-n+1}^{y}T_i
\end{equation}

## Sensitivity Analyses {#app:HBMFits-2}

Some additional model runs were implemented to evaluate some different features of the HBM.  See Table \@ref(tab:SensitivityRuns) for brief descriptions of these additional model runs. We term the model run using above described specifications for the HBM the “base case”.  Note that the results reported below were obtained from earlier WinBUGS 1.4.3 code versions of the HBM which included a simpler approximation for computing Smsy [i.e., from Table 7.2 of @HilbornWalters1992] than was used in the main body of the report.  The results from these additional runs are summarized further below.   


  
  <!-- TABLE 2. Sensitivity Analyses -->
```{r SensitivityRuns, echo=FALSE, results='asis'}
cap <- "Description of sensitivity runs to evalute the sensitivity of estimation results to model structure and some key inputs."

data.frame(
  `Run` = c(seq(1,6), "7-24", "25"), 
  Description = c(
    "HBM base case but including Korman and English’s (2013) coding in computing precision in the likelihood function from $\\sigma$",
    "Same as HBM base case but with no prior lower bounds on $\\beta$.",
    "Same as HBM base case but leaving out common shared year effects.",
    "Non-hierarchical model run with no common shared year effect but including the same $S_{\\textrm{max}}$ prior information as in the base case HBM.",
    "Same as HBM base case but normal priors on $S_{\\textrm{max}}$ instead of the base case lognormal prior on Ricker $\\beta$.",
    "Same as HBM base case but with vague Ricker $\\beta$ priors, but including the prior lower bound on Ricker $\\beta$.",
    "Leave out stock-recruit data in the HBM, one stock at a time.",
    "Leave out stock-recruit data for the Babine Enhanced stocks in the HBM, i.e., Pinkut and Fulton."
  ),
  check.names = FALSE
) %>%
  csas_table(
    format = "latex", 
    escape = FALSE, 
    font_size = 10, 
    # align = c("p{1.2cm}","p{12cm}"),
    align = c("l", "l"),
    caption = cap, #paste("(tab:SensitivityRuns)", cap)
  ) %>%
  kableExtra::row_spec(seq(1,7), hline_after = TRUE) %>%
  kableExtra::column_spec(2, width = "12cm")
```


## Results

### Shrinkage in the Ricker $a$ parameter in going from a non-HBM to an HBM

```{r, echo=FALSE}
shrink <- (HBM.results$Shrinkage$Shrinkage * 100) %>% range() %>% round
kitwanga <- round((HBM.results$Shrinkage %>% filter(Stock == "Kitwanga") %>% select(Shrinkage) %>% unlist)*100 )
pinkut <- round((HBM.results$Shrinkage %>% filter(Stock == "Pinkut") %>% select(Shrinkage) %>% unlist)*100 )
fulton <- round((HBM.results$Shrinkage %>% filter(Stock == "Fulton") %>% select(Shrinkage) %>% unlist)*100 )
other <- HBM.results$Shrinkage %>% filter(Stock %in% c("Kitwanga", "Pinkut", "Fulton" == FALSE)) %>% select(Shrinkage) %>% unlist %>% range
other <-round(other*100,1)
sig.sd <- (HBM.results$`Shrinkage`$Diff.SD * -100) %>% range() %>% round(digits = 1)  # % reduction
```

In both the non-HBM and HBM models a common shared year effect was estimated to make the shrinkage analysis valid.  Under the non-HBM the prior for the Ricker $a$ parameter for each stock was similarly lognormal, had $a$ prior median of $\log(0.5)$ and precision of 1.  Under the non-HBM a moderate amount of shrinkage can be seen (Figure \@ref(fig:HBMShrinkage)). For example the Asitka stock which had the highest posterior median value for the Ricker $a$ parameter under both models showed a `r -shrink[1]`% decrease in the Ricker $a$ parameter (Table Z.Z).  This was the largest percentage decrease among the Skeena sockeye salmon stocks included in the HBM.  In contrast the Kitwanga stock which had the second lowest posterior median for the Ricker $a$ parameter under the non-HBM showed the largest increase of `r kitwanga`% when going from the non-HBM to the HBM.  In contrast the Pinkut and Fulton stocks which had posterior medians for Ricker a under the non-HBM close to the middle range of the Posterior median estimates showed very little shrinkage, i.e., `r pinkut`% and `r fulton`%, respectively (Figure \@ref(fig:HBMShrinkage), Table Z.Z).  For all of the 18 Skeena sockeye salmon stocks the posterior SDs for the Ricker $a$ parameter were all smaller, i.e., `r sig.sd[1]`% to `r sig.sd[2]`% smaller, under the HBM than under the non-HBM (Table Z.Z).



```{r, echo=FALSE}
cap <- "Shrinkage plot for posterior median values for the Ricker $a$ parameter obtained under non-HBM and HBM models of stock-recruit data for 18 Skeena River sockeye salmon stocks."
```
```{r HBMShrinkage, echo=FALSE, fig.width=7, fig.cap=cap}
HBM.results$Shrinkage %>%
  select(Stock, nonHBM.median, HBM.median) %>%
  gather(key=Model, value = Median, ends_with("median")) %>%
  mutate(
    Model = str_replace(Model, "\\.median", "") %>% factor(levels = c("nonHBM", "HBM"))
  ) %>%
  ggplot(., aes(x=Model, y = Median)) +
  geom_point(aes(color = Stock)) +
  geom_line(aes(group = Stock, color = Stock)) +
  theme_classic(14) +
  theme(legend.title = element_blank()) +
  labs(
    y = "Posterior Median"
  )
```


```{r HBMShrinkageVal, echo=FALSE, results='asis'}
cap <- "Posterior median estimates and standard deviations (SD) for the Ricker $a$ parameter for the 18 Skeena River sockeye salmon stocks under non-HBM and HBM models  The percentage change shows the percentage change in the parameter estimate in going from the non-HBM to the HBM implementation.  "

HBM.results[['Shrinkage']] %>%
  mutate(
    Shrinkage = paste0(round(Shrinkage * 100,1), "\\%"),
    Diff.SD = paste0(round(Diff.SD * 100,1), "\\%")
  ) %>%
  rename(
    `nonHBM Median` = nonHBM.median,
    `HBM Median` = HBM.median,
    `nonHBM SD` = nonHBM.SD,
    `HBM SD` = HBM.SD,
    `SD \\% Diff` = Diff.SD,
  ) %>%
  csas_table(
    format = "latex",
    escape = FALSE, 
    font_size = 10,
    align = c("l", rep("r", 6)),
    caption = cap #paste("(ref:tabSenRun1)", cap)
  ) %>%
  kableExtra::column_spec(2:7, width = "1.5cm")
```

\newpage
### Sensitivity Run 1:  Effect of Korman and English’s (2013) coding error for their HBM. {#app:HBMFits-2-1}
```{r, echo=FALSE}
sig.mean <- (HBM.results$`Run 1`$Diff.mean * 100) %>% range() %>% round
sig.sd <- (HBM.results$`Run 1`$Diff.SD * 100) %>% range() %>% round
```


There was a coding error in the model used by @KormanEnglish2013; their code incorrectly transformed the parameter $\sigma_i$ to $\tau_i$ using the incorrect code:  \texttt{tau[i]<-pow(sd[i],-0.5)} which instead should have been \texttt{tau[i]<-pow(sd[i],-2)}. This coding error changed very little the HBM posterior results for the Ricker $a$ and $\beta$ parameters for the 18 stocks (Table \@ref(tab:SenRun1)).  However, the estimates of the $\sigma_i$ parameter were between `r sig.mean[1]`% and `r sig.mean[2]`%  of those obtained under the HBM base case.   The posterior SDs for $\sigma_i$ ranged between `r sig.sd[1]`% and `r sig.sd[2]`%  larger than those obtained under the base case.  Either the lower or upper boundary point on the prior for $\sigma_i$ was hit for some of the stocks.    If the coding error had persisted, any simulations of stock-recruit data using the estimated values for $\sigma_i$ may have led to results with excessive variability.  It is thus essential for this coding error to be corrected in any further implementations of Korman and English’s (2013) HBM.

<!-- TABLE Sensitivity Run 1 -->
```{r SenRun1, echo=FALSE, results='asis'}
cap <- "Posterior means and posterior standard deviations for $\\sigma$ from the base case (coding error removed) and model run that included Korman and English’s (2013) coding error."

HBM.results[['Run 1']] %>%
  mutate(
    Diff.mean = paste0(round(Diff.mean*100), "\\%"),
    Diff.SD = paste0(round(Diff.SD * 100), "\\%")
  ) %>%
  rename(
    `Base Case mean $\\sigma$` = Base.Case.mean,
    `Code Error mean $\\sigma$` = Code.Error.mean,
    `\\% Diff` = Diff.mean,
    `Base Case SD($\\sigma$)` = Base.Case.SD,
    `Code Error SD($\\sigma$)` = Code.Error.SD,
    `\\% Diff ` = Diff.SD,
  ) %>%
  csas_table(
    format = "latex",
    escape = FALSE, 
    font_size = 10,
    align = c("l", rep("r", 6)),
    caption = cap #paste("(ref:tabSenRun1)", cap)
  ) %>%
  kableExtra::column_spec(2:7, width = "1.5cm")
```




### Sensitivity Run 2:  Same as HBM base case but with no prior lower bound on the Ricker $\beta$ parameter



When there was no prior lower bound included the Ricker $\beta$ parameter by stock, the estimates of $S_\textrm{max}$ and MSY-based reference points included some extremely high MCMC chain values that were way out in the tails of the Markov chains and very far removed from the range of values with support from the data; for these stocks, posterior means, medians, and probability interval values were highly sensitive to the inclusion of these extreme outlier values in the chains (see Table \@ref(tab:SenRun2) for some example results for $S_\textrm{MSY}$).  It is common in implementations of MCMC implementations such as WinBUGS to set bounds on key variables in the model to prevent extreme outlier values in the chains from affecting the posterior calculations [e.g., @MeyerMillar1999; @MichielsensMcAllister2004].  It is thus recommended that prior lower bounds on the Ricker $\beta$ parameter be implemented for each stock to eliminate this source of bias.


<!-- TABLE X.A Sensitivity Run 2 -->
```{r SenRun2, echo=FALSE, results='asis'}
cap <- "Posterior means and posterior standard deviations for Smsy from the base case and model run where no prior upper bounds were placed on Smax."

HBM.results[['Run 2']] %>%
  mutate(
    Base.Case.mean = round(Base.Case.mean),
    No.Bound.mean = round(No.Bound.mean),
    Base.Case.SD = round(Base.Case.SD),
    No.Bound.SD = round(No.Bound.SD),
    Diff.mean = paste0(Diff.mean*100, "\\%"),
    Diff.SD = paste0(Diff.SD * 100, "\\%")
  ) %>%
  rename(
    `Base Case mean $S_\\textrm{MSY}$` = Base.Case.mean,
    `No Bound mean $S_\\textrm{MSY}$` = No.Bound.mean,
    `\\% Diff` = Diff.mean,
    `Base Case SD($S_\\textrm{MSY}$)` = Base.Case.SD,
    `No Bound SD($S_\\textrm{MSY}$)` = No.Bound.SD,
    `\\% Diff ` = Diff.SD
  ) %>%
  csas_table(
    format = "latex",
    escape = FALSE, 
    font_size = 10,
    align = c("l", rep("r", 6)),
    caption = cap, #paste("(ref:tabSenRun2)", cap)
  ) %>%
  kableExtra::column_spec(2:7, width = "1.5cm")
```

### Sensitivity Run 3: Same as HBM base case but leaving out common shared year effects

```{r, echo=FALSE}
sig.mean <- (HBM.results[['Run 3']]$Diff.mean * 100)  %>% range() %>% round
sig.sd <- (HBM.results[['Run 3']]$Diff.SD * 100) %>% range()  %>% round
```

Significant, strong common shared year effects were estimated for numerous years in the 1960-2014 brood year time series.  Separating out this effect provided more precise estimates of Ricker stock recruit parameters for several of the stocks and more precise estimates of management quantities of interest.  This also allowed for common shared systematic change in stock productivity to be estimated.  The posterior means for the Smsy reference point by stock for example were between about `r sig.mean[1]`% and `r sig.mean[2]`% different between runs when common shared year effects were excluded versus accounted for (Table \@ref(tab:SenRun3)).  When no common shared year effect was included in the HBM, estimates of Ricker stock-recruit parameters and associated management parameters for several of the stocks were on average less precisely estimated with posterior SDs being up to about `r sig.sd[2]`% larger for $S_\textrm{msy}$ for several of the stocks.

<!-- TABLE X.B Sensitivity Run 2 -->
```{r SenRun3, echo=FALSE, results='asis'}
cap <- "Posterior means and posterior standard deviations for Smsy from the base case and model run where no prior upper bounds were placed on Smax."

HBM.results[['Run 3']] %>%
  mutate(
    Base.Case.mean = round(Base.Case.mean),
    No.TE.mean = round(No.TE.mean),
    Base.Case.SD = round(Base.Case.SD),
    No.TE.SD = round(No.TE.SD),
    Diff.mean = paste0(Diff.mean*100, "\\%"),
    Diff.SD = paste0(Diff.SD * 100, "\\%")
  ) %>%
  rename(
  	`Base Case mean $S_\\textrm{MSY}$` = Base.Case.mean,
  	`No \\newline $T_y$ \\newline mean $S_\\textrm{MSY}$` = No.TE.mean,
  	`\\% Diff` = Diff.mean,
  	`Base Case SD($S_\\textrm{MSY}$)` = Base.Case.SD,
  	`No \\newline $T_y$ \\newline SD($S_\\textrm{MSY}$)` = No.TE.SD,
  	`\\% Diff ` = Diff.SD
  ) %>%
  csas_table(
    format = "latex",
    escape = FALSE, 
    font_size = 10,
    align = c("l", rep("r", 6)),
    caption = cap #paste("(ref:tabSenRun3)", cap)
  ) %>%
  kableExtra::column_spec(2:7, width = "1.5cm")
```


### Sensitivity Run 4: Non-hierarchical model run with no common shared year effect but including the same $S_\textrm{max}$ prior information as in the base case HBM

```{r, echo=FALSE}
sig.mean <- (HBM.results[['Run 4']]$Diff.mean * 100) %>% range()  %>% round
sig.sd <- (HBM.results[['Run 4']]$Diff.SD * 100) %>% range()  %>% round

smsy.diff <- filter(HBM.results[['Run 4']], Stock == "Sum Smsy across stocks") %>% select(Diff.mean, Diff.SD) %>% unlist * 100 %>% round
```


When a nonhierarchical Bayesian model(nonHBM) with no common shared year effects was run, estimates of Ricker stock-recruit parameters and associated management parameters for several of the stocks were on average less precisely estimated with posterior SD in Smsy for example on average being about `r sig.sd[1]`% to `r sig.sd[2]`% different between the nonHBM and HBM runs (Table \@ref(tab:SenRun4)).  Percentage differences between the nonHBM and HBM for posterior mean estimates for Smsy parameters ranged by stock between `r sig.mean[1]`% and `r sig.mean[2]`%.  These results indicate that on average Ricker and msy-based parameter estimates are more precisely estimated with the HBM and results for some quantities for some stocks can differ considerably.  However, the sum of the $S_\textrm{MSY}$ estimates across stocks was only about `r smsy.diff['Diff.mean']`% larger under the non-hierarchical model but the posterior SD in the $S_\textrm{MSY}$ values summed over the Skeena River stocks was `r smsy.diff['Diff.SD']`% larger than that for the HBM (Table \@ref(tab:SenRun4)).

<!-- TABLE X.C Sensitivity Run 4 -->
```{r SenRun4, echo=FALSE, results='asis'}
cap <- "Posterior means and posterior standard deviations for $S_\\textrm{MSY}$ from the base case HBM and nonHBM run."

HBM.results[['Run 4']] %>%
  mutate(
    Base.Case.mean = round(Base.Case.mean),
    nonHBM.mean = round(nonHBM.mean),
    Base.Case.SD = round(Base.Case.SD),
    nonHBM.SD = round(nonHBM.SD),
    Diff.mean = paste0(Diff.mean*100, "\\%"),
    Diff.SD = paste0(Diff.SD * 100, "\\%")
  ) %>%
  rename(
  	`Base Case mean $S_\\textrm{MSY}$` = Base.Case.mean,
  	`nonHBM \\newline mean $S_\\textrm{MSY}$` = nonHBM.mean,
  	`\\% Diff` = Diff.mean,
  	`Base Case SD($S_\\textrm{MSY}$)` = Base.Case.SD,
  	`nonHBM \\newline SD($S_\\textrm{MSY}$)` = nonHBM.SD,
  	`\\% Diff ` = Diff.SD
  ) %>%
  csas_table(
    format = "latex",
    escape = FALSE, 
    font_size = 10,
    align = c("l", rep("r", 6)),
    caption = cap 
  ) %>%
  kableExtra::column_spec(2:7, width = "1.5cm")
```

### Sensitivity Run 5: Normal priors on Smax in the HBM instead of the lognormal prior on Ricker $\beta$

```{r, echo=FALSE}
sig.mean <- (HBM.results[['Run 5']]$Diff.mean * 100) %>% range()  %>% round
sig.sd <- (HBM.results[['Run 5']]$Diff.SD * 100) %>% range()  %>% round
```

When a normal prior for $S_\textrm{max}$ was applied instead of a lognormal prior on the Ricker $\beta$ parameter in the HBM, but otherwise using the same input information from the lake productivity analyses, the posterior estimates for several of the quantities became less precise and posterior estimates differed markedly for some of the quantities.  Posterior SDs for $S_\textrm{MSY}$ were for example much larger on average, e.g., between about `r sig.sd[1]`% and `r sig.sd[2]`% larger (Table \@ref(tab:SenRun5)).  Posterior mean estimates for $S_\textrm{MSY}$ differed between the two model runs by `r sig.mean[1]`% to `r sig.mean[2]`%.  Though it appears the same information is used in a normal prior for $S_\textrm{max}$, this prior on average loses information about the Ricker $\beta$ parameter compared to a prior for the Ricker $\beta$ parameter that uses the same $S_\textrm{max}$ information. 

```{r SenRun5, echo=FALSE, results='asis'}
cap <- "Posterior means and posterior standard deviations for $S_\\textrm{MSY}$ from the base case and model run with a Normal priors on $S_\\textrm{max}$."

HBM.results[['Run 5']] %>%
  mutate(
    Base.Case.mean = round(Base.Case.mean),
    Normal.Prior.mean = round(Normal.Prior.mean),
    Base.Case.SD = round(Base.Case.SD),
    Normal.Prior.SD = round(Normal.Prior.SD),
    Diff.mean = paste0(Diff.mean*100, "\\%"),
    Diff.SD = paste0(Diff.SD * 100, "\\%")
  ) %>%
  rename(
  	`Base Case mean $S_\\textrm{MSY}$` = Base.Case.mean,
  	`Normal Prior mean $S_\\textrm{MSY}$` = Normal.Prior.mean,
  	`\\% Diff` = Diff.mean,
  	`Base Case SD($S_\\textrm{MSY}$)` = Base.Case.SD,
  	`Normal Prior SD($S_\\textrm{MSY}$)` = Normal.Prior.SD,
  	`\\% Diff ` = Diff.SD
  )  %>%
  csas_table(
    format = "latex",
    escape = FALSE, 
    font_size = 10,
    align = c("l", rep("r", 6)),
    caption = cap 
  ) %>%
  kableExtra::column_spec(2:7, width = "1.5cm")
```


### Sensitivity Run 6: Application of vague Ricker $\beta$ priors in the HBM, but including the prior lower bound on $\beta$
```{r, echo=FALSE}
sig.mean <- (HBM.results[['Run 6']]$Diff.mean * 100) %>% range() %>% round
sig.sd <- (HBM.results[['Run 6']]$Diff.SD * 100) %>% range() %>% round
```

When a vague prior was applied for the Ricker $\beta$ prior for all 18 Skeena sockeye salmon stocks in the HBM, posterior estimates for the abundance-based management quantities were less precise on average with posterior SDs, e.g., for $S_\textrm{MSY}$ ranging between `r sig.sd[1]`% to `r sig.sd[2]`% larger under the vague priors (Table \@ref(tab:SenRun6)).  Posterior means for $S_\textrm{MSY}$ estimates in the HBM that used vague priors for the Ricker $\beta$ parameter were between about `r sig.mean[1]`% and `r sig.mean[2]`% different from those obtained under the mixture of informed and vague priors in the base case HBM.  The use of informative priors for the Ricker $\beta$ parameter based on prior information on $S_\textrm{max}$ via the lake productivity analyses thus combined with the data to provide more precise stock-recruit parameter estimates that in some cases differed from the less precise estimates given by the stock-recruit data.  

```{r SenRun6, echo=FALSE, results='asis'}
cap <- "Posterior means and posterior standard deviations for $S_\\textrm{MSY}$ from the base case and model run with a Normal priors on $S_\\textrm{max}$."

HBM.results[['Run 6']] %>%
  mutate(
    Base.Case.mean = round(Base.Case.mean),
    Vague.Prior.mean = round(Vague.Prior.mean),
    Base.Case.SD = round(Base.Case.SD),
    Vague.Prior.SD = round(Vague.Prior.SD),
    Diff.mean = paste0(Diff.mean*100, "\\%"),
    Diff.SD = paste0(Diff.SD * 100, "\\%")
  ) %>%
  rename(
  	`Base Case mean $S_\\textrm{MSY}$` = Base.Case.mean,
  	`Vague $\\beta$ Prior mean $S_\\textrm{MSY}$` = Vague.Prior.mean,
  	`\\% Diff` = Diff.mean,
  	`Base Case SD($S_\\textrm{MSY}$)` = Base.Case.SD,
  	`Vague $\\beta$ Prior SD($S_\\textrm{MSY}$)` = Vague.Prior.SD,
  	`\\% Diff ` = Diff.SD
  )  %>%
  csas_table(
    format = "latex",
    escape = FALSE, 
    font_size = 10,
    align = c("l", rep("r", 6)),
    caption = cap 
  ) %>%
  kableExtra::column_spec(2:7, width = "1.5cm")
```


### Sensitivity Runs 7-25.  Leave out stock-recruit data in the HBM, one stock at a time

To evaluate the relative influence of each stock-recruit data set from each stock, the HBM was run dropping out the stock-recruit data for one stock at a time.  In run 25 the stock-recruit data for both of the Babine Enhanced stocks (i.e., Fulton and Pinkut) were left out of the HBM.  The posterior predictive distribution for the Ricker $a$ parameter for an “unsampled” stock was computed for each of these runs (see equation \@ref(eq:unsamp)).  The posterior predictive distributions were plotted under each of the HBM runs 7-25 and under the base case HBM shows that the posterior predictive distributions were very similar between all HBM runs 7-25 and the base case HBM run (Figure \@ref(fig:jackProd)).  The posterior mean estimates of the time series of common shared year effects also did not change their sign or markedly change in magnitude when the stock-recruit data from one of the stocks or both of the Babine enhanced stocks were removed from the HBM (Figure \@ref(fig:jackTE)).  These results indicate that the stock-recruit data from no one stock and from both of the enhanced Babine sockeye salmon stocks had a noticeable influence on the HBM results.  

```{r, echo=FALSE}
cap <- "Posterior predictive distributions for the Ricker $a$ parameter when the sockeye salmon stock-recruit data from one stock at a time was dropped from the HBM for the Skeena watershed (runs 7-24) and the stock-recruit data were dropped for both of the Babine Enhanced sockeye salmon stocks, i.e., both Pinkut and Fulton (run 25). Base Case posterior predictive distribution is coloured dark grey, with colours used for runs 7-25."
```
```{r jackProd,   fig.cap=cap}
include_graphics("data/HBM/Appendix-HBM-Productivity-jackknife.pdf")
```


```{r, echo=FALSE}
cap <- "Posterior mean estimates for the common shared year effect when the sockeye salmon stock-recruit data from one stock at a time was dropped from the HBM for the Skeena watershed (runs 7-24) and the stock-recruit data were dropped for both of the Babine Enhanced sockeye salmon stocks, i.e., both Pinkut and Fulton (run 25). Base Case posterior predictive distribution is coloured dark grey, with colours used for runs 7-25."
```
```{r jackTE,   fig.cap=cap}
include_graphics("data/HBM/Appendix-HBM-yeareffect-jackknife.pdf")
```